Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-12 linear_adia.tex
generalized bcrit (attempt 1)

Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
simultaneous equations for $\hat{\omega}$ and the critical thermal
timescale $\beta$ for marginal stability. If $\beta\ll 1$,
then a simple solution to Eq. \ref{relax_cond_simp1} is 
\begin{align}
  \left(\hat{\omega}\khat\right)^2 \simeq M+1, 
\end{align}
which is the dispersion relation for low-frequency waves in vertically
isothermal disks without vertical shear (see Appendix
\ref{stable_novshear}). Then  
\begin{align}\label{bcrit_gen}
  \beta = \frac{\epsilon|q|}{(\gamma - 1)}
  \frac{(1+M)^{3/2}}{(1+2M)^2} \leq \beta_\mathrm{crit} 
\end{align}
where 
\begin{align}\label{iso_vsi_cond}
  \beta_\mathrm{crit}  \equiv \frac{\epsilon|q|}{\gamma-1}
\end{align}
is the critical thermal timescale for the fundamental mode
($M=0$). However, according to Eq. \ref{bcrit_gen}  the critical
thermal timescale decreases with increasing  $M$. We thus expect
higher order modes to be stabilized more rapidly by finite thermal
relaxation $\beta$. When $\beta>\beta_\mathrm{crit}$, all modes
(satisfying the above assumptions) are stabilized. We can interpret
$\beta_\mathrm{crit}$ as the thermal timescale below which
perturbations are effectively isothermal to allow the VSI.  We will confirm
Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.

(We remark that the critical thermal timescale $\beta_\mathrm{crit}$
for the fundamental mode may be obtained more readily with fewer
assumptions by setting $M=0$ in
Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset. )

>>>>

Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
simultaneous equations for $\hat{\omega}$ and the critical thermal
timescale $\beta$ for marginal stability. If $\beta\ll 1$,
then a simple solution to Eq. \ref{relax_cond_simp1} is 
\begin{align}
  \left(\hat{\omega}\khat\right)^2 \simeq M+1, 
\end{align}
which is the dispersion relation for low-frequency waves in vertically
isothermal disks without vertical shear (see Appendix
\ref{stable_novshear}). Then  
\begin{align}\label{bcrit_gen}
  \beta = \frac{\epsilon|q|}{(\gamma - 1)}
  \frac{(1+M)^{3/2}}{(1+2M)^2} \leq \beta_\mathrm{crit} 
\end{align}
where 
\begin{align}\label{iso_vsi_cond}
  \beta_\mathrm{crit}  \equiv \frac{\epsilon|q|}{\gamma-1}
\end{align}
is the critical thermal timescale for the fundamental mode
($M=0$). However, according to Eq. \ref{bcrit_gen}  the critical
thermal timescale decreases with increasing  $M$. We thus expect
higher order modes to be stabilized more rapidly by finite thermal
relaxation. When $\beta>\beta_\mathrm{crit}$, all modes
(satisfying the above assumptions) are stabilized. We can interpret
$\beta_\mathrm{crit}$ as the thermal timescale below which
perturbations are effectively isothermal to allow the VSI.  We will confirm
Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.

(We remark that the critical thermal timescale $\beta_\mathrm{crit}$
for the fundamental mode may be obtained more readily with fewer
assumptions by setting $M=0$ in
Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset.)
